---
title: 'Class 11: Scraping Multiple Webpages'
author: "Isobel Bartels"
date: "2026-09-30"
output: 
  html_document:
    toc: true
    toc_float: true
---


##Learning objectives
- Enable you to scrape multiple webpages (from the same site), using a 'for' loop type approach 

- Enable you to scrape multiple webpages from the same site, using the [purrr package] (https://purrr.tityverse.org/), which is a functional programming approach 
*Note* in this process, we will create our own function

## Re-examining the bullentin example 
let us try to expand on the example from last class to include the following:
[ACC](https://bulletin.miamioh.edu/courses-instruction/acc/),
[BUS](https://bulletin.miamioh.edu/courses-instruction/bus/),
[ISA](https://bulletin.miamioh.edu/courses-instruction/isa/), and
[MKT](https://bulletin.miamioh.edu/courses-instruction/mkt/).

```{r bulletin}
dept_abbrv = c('acc','bus','isa','mkt')

courses_df = data.frame() #initializing a container to store stuff from the loop
for (abbrv in dept_abbrv) {
  cat( rep("-", 50) , "\n")
  print(abbrv)
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
  )
  
  #Q: why did fadel make this an R object instead of a strong like last class? 
  #[A] we did this so we can resure the rest of our code with only changing the URl 
  webpage = rvest::read_html(bulletin_url) 
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  print(df$title[1:2])
  courses_df = rbind(courses_df, df)
}


```

## 'purr' approach for the same problem 
this is a functional programming approach, we will write our own function. In R, we use a function called 'fuction' to create a function, as opposed to python, R functions return a single object 

```{r purr}
dept_abbrv = c('acc','bus','isa','mkt')
scrape_mu_bulletin = function(abbrv) {
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
  )
  
  #Q: why did fadel make this an R object instead of a strong like last class? 
  #[A] we did this so we can resure the rest of our code with only changing the URl 
  webpage = rvest::read_html(bulletin_url) 
  
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> titles
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  return(df)
}
scrape_mu_bulletin = function(abbrv)
test = scrape_mu_bulletin(abbrv = 'eco')
#to use functions from purrr, we will use a map_*(), or map2_*(), or a pmap_*()
#map uses a function that takes a single input( or you are only changng a single input)

#map2 uses a function that takes two inputs 
#pmap >=3
# after the _, you put the type of object you are getting back


```

#Scrape data from the website planecrashinfo
the website that we 
```{r}
year = 2025

crash_url = paste0('https://www.planecrashinfo.com/',year, )
```

