---
title: "Midwest Airbnb: Eight Analyses"
author: "Isobel Bartels"
date: "2026-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb_extended.db")
listings = DBI::dbReadTable(con, "listings") |>
  tibble::as_tibble()
availability = DBI::dbReadTable(con, "availability_monthly") |>
  tibble::as_tibble()
reviews = DBI::dbReadTable(con, "reviews_monthly") |>
  tibble::as_tibble()
DBI::dbDisconnect(con)
```

## 1. Who owns the market

Host size is counted across all three cities before calculating Chicago's share.

```{r market-ownership}
host_counts = listings |> dplyr::count(host_id, name = "host_listing_count")
listings_with_host_counts = listings |>
  dplyr::left_join(host_counts, by = "host_id")
A1 = sum(host_counts$host_listing_count >= 10)
chicago_share = listings_with_host_counts |>
  dplyr::filter(city == "Chicago") |>
  dplyr::summarise(percent = 100 * mean(host_listing_count >= 10))
A2 = chicago_share$percent
A1
A2
```

## 2. Value for a family of four

```{r family-value}
family_value = listings |>
  dplyr::filter(room_type == "Entire home/apt", accommodates >= 4,
         number_of_reviews >= 10, review_scores_rating >= 4.8) |>
  dplyr::mutate(price_per_person = price / accommodates) |>
  dplyr::group_by(city, neighbourhood) |>
  dplyr::summarise(n_listings = dplyr::n(),
            median_price_per_person = median(price_per_person, na.rm = TRUE),
            .groups = "drop") |>
  dplyr::filter(n_listings >= 20)
columbus_best_value = family_value |>
  dplyr::filter(city == "Columbus") |>
  dplyr::filter(median_price_per_person == min(median_price_per_person, na.rm = TRUE))
A3 = columbus_best_value$neighbourhood
columbus_best_value
A3
```

## 3. What earns

```{r earnings}
earnings = listings |>
  dplyr::group_by(city, room_type) |>
  dplyr::summarise(dplyr::across(c(price, reviews_per_month, estimated_revenue_l365d),
                   ~ median(.x, na.rm = TRUE)), .groups = "drop")
twin_cities_earnings = earnings |>
  dplyr::filter(city == "Twin Cities", room_type == "Entire home/apt")
A4 = twin_cities_earnings$estimated_revenue_l365d
earnings
A4
```

## 4. Concentration

Exactly `ceiling(0.10 * dplyr::n())` rows per city are marked; original row order
breaks ties at the cutoff. Missing revenue is skipped when summing.

```{r concentration}
revenue_ranked = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(estimated_revenue_l365d), .by_group = TRUE) |>
  dplyr::mutate(top_ten_percent = dplyr::row_number() <= ceiling(0.10 * dplyr::n())) |>
  dplyr::ungroup()
concentration = revenue_ranked |>
  dplyr::group_by(city) |>
  dplyr::summarise(top_revenue = sum(estimated_revenue_l365d[top_ten_percent], na.rm = TRUE),
            total_revenue = sum(estimated_revenue_l365d, na.rm = TRUE),
            percent = 100 * top_revenue / total_revenue, .groups = "drop")
columbus_concentration = concentration |> dplyr::filter(city == "Columbus")
A5 = columbus_concentration$percent
concentration
A5
```

## 5. The year ahead

The denominator is entire homes with an October 2026 availability row.
Share is reported as a proportion (0 to 1), with a percentage alongside it.

```{r year-ahead}
october_entire_homes = listings |>
  dplyr::filter(room_type == "Entire home/apt") |>
  dplyr::inner_join(availability |> dplyr::filter(month == "2026-10"), by = "id") |>
  dplyr::mutate(mostly_taken = nights_available < 0.5 * nights_in_month)
october_shares = october_entire_homes |>
  dplyr::group_by(city) |>
  dplyr::summarise(n_listings = dplyr::n(), share = mean(mostly_taken), .groups = "drop") |>
  dplyr::mutate(percent = 100 * share)
lowest_share = october_shares |> dplyr::filter(share == min(share))
A6 = lowest_share$city
A7 = lowest_share$share
october_shares
A6
A7
```

## 6. Demand by season

Monthly review counts are summed, rather than counting monthly table rows.
The date range includes July 2025 and June 2026.

```{r seasonality}
monthly_demand = reviews |>
  dplyr::inner_join(listings |> dplyr::select(id, city), by = "id") |>
  dplyr::filter(month >= "2025-07", month <= "2026-06") |>
  dplyr::group_by(city, month) |>
  dplyr::summarise(total_reviews = sum(reviews, na.rm = TRUE), .groups = "drop")
twin_cities_demand = monthly_demand |> dplyr::filter(city == "Twin Cities")
busiest_month = twin_cities_demand |>
  dplyr::filter(total_reviews == max(total_reviews))
A8 = busiest_month$month
A9 = max(twin_cities_demand$total_reviews) / min(twin_cities_demand$total_reviews)
twin_cities_demand
A8
A9
```

## 7. The Superhost premium

The premium is a dollar difference between medians, not a percentage.

```{r superhost-premium}
superhost_summary = listings |>
  dplyr::filter(room_type == "Entire home/apt", (host_is_superhost == "t" | host_is_superhost == "f")) |>
  dplyr::group_by(city, host_is_superhost) |>
  dplyr::summarise(median_price = median(price, na.rm = TRUE),
            median_rating = median(review_scores_rating, na.rm = TRUE),
            .groups = "drop")
B1 = superhost_summary |>
  dplyr::group_by(city) |>
  dplyr::summarise(premium_dollars = median_price[host_is_superhost == "t"] -
              median_price[host_is_superhost == "f"], .groups = "drop")
superhost_summary
print(as.data.frame(B1))
```

## 8. What the join lost

This checks for no availability rows in any month, not just October.

```{r missing-availability}
missing_availability = listings |> dplyr::anti_join(availability, by = "id")
missing_by_city = missing_availability |> dplyr::count(city, sort = TRUE)
A10 = nrow(missing_availability)
A10
missing_by_city
```

## B1. Advice to a Twin Cities host

Premiums: **Chicago $43.75; Columbus $40.50; Twin Cities -$1.38.**

The Twin Cities Superhost median is $1.38 lower than the other-host median, so
these data do not suggest that the badge alone will let you charge more.
This comparison does not establish what earning the badge would do to your own
price, and I would compare reviews per month before deciding it is not worth
pursuing, since Superhosts could attract more demand even at similar prices.

## B2. Market concentration and a ten-listing limit

```{r city-concentration}
city_concentration = listings_with_host_counts |>
  dplyr::group_by(city) |>
  dplyr::summarise(listing_share = 100 * mean(host_listing_count >= 10)) |>
  dplyr::left_join(
    concentration |> dplyr::select(city, revenue_share = percent),
    by = "city"
  )
print(as.data.frame(city_concentration))
```

Columbus would likely be changed most by a ten-listing cap because hosts with
10 or more listings hold 37.15% of its listings, compared with 31.90% in Chicago
and 17.16% in the Twin Cities; the top 10% of listings earn 36.08%, 40.74%, and
40.27% of estimated revenue in those cities, respectively.
These shares suggest exposure to the rule rather than its exact effect, since
hosts with exactly ten listings would already comply.
The data cannot tell the council whether multiple host IDs belong to the same
business, so host counts may not represent business ownership.
